% !TeX root = forth.tex
% !TeX spellcheck = en_US

\chapter*{200\emph{x} Membership}
\label{members}
\addcontentsline{toc}{section}{200\emph{x} Membership}
\markboth{200\emph{x} Membership}{200\emph{x} Membership}

This document is maintained by the Forth 200\emph{x} Standards
Committee. The committee meetings are open to the public, anybody
is allowed to \replace{x:membership}{join the committee in its deliberations}{
attend a physical meeting as an observer}.

\remove{x:membership}{%
Membership of the committee is open to anybody who can attend.  On
attending a meeting of any kind a non-member becomes an observing
member (observer).  If they attend the next voting meeting, they
will become a voting member of the committee, otherwise they revert
to non-member status.  An observer will not normally be allowed to
vote, but may be allowed at the discretion of the committee.
A member will be deemed to have resigned from the committee if they
fail to attend two consecutive voting meetings.}

\place{x:membership}{%
Committee membership is open to anybody who can attend physical
meetings.  On attending a physical meeting, a non-member becomes
an observer.  An observer becomes a voting member by attending the
next physical meeting.  An observer will not normally be allowed to
vote except at the discretion of the committee.  A member will be
deemed to have resigned from the committee by failing to attend two
consecutive physical meetings.}

\place{x:membership}{%
Electronic meetings of the committee are held between physical
meetings as required.}

Currently the committee has the following voting members:

\def\shortyear#1{\expandafter\shortyearhelper#1}
\def\shortyearhelper#1#2#3#4{#3#4}
\newsavebox{\emailbox}

\newcommand{\member}[6][\empty]{% [+/-]{name}{domain}{user}{Company}{City}
	\savebox{\emailbox}{\texttt{\randomize{#4@#3}}}
	\let\mod=\relax
	\ifx+#1\let\mod=\uline\fi
	\ifx-#1\let\mod=\sout\fi

	\makebox[0.5\linewidth][l]{\mod{#2}\dotfill} \quad \makebox{\mod{#5}}
	\ifx\empty#1\else\cbstart\patch{ed\shortyear{\the\year}}\fi \\
	\makebox[0.5\linewidth][l]{\mod{\usebox{\emailbox}}} \quad \mod{#6}\\[-\baselineskip]
	\ifx\empty#1\else\cbend\fi \par
}

\html{<table class="center" width="600px">}
\member{Willem Botha}
	{ccssa.com}{willem.botha}
	{Construction Computer Software (Pty) Ltd}
	{Cape Town, South Africa}

\member{Dr.\ M.\ Anton Ertl (Chair)}
	{mips.complang.tuwien.at}{anton}
	{Technische Universit\"at Wien}
	{Wien, Austria}

\member[-]{Andrew Haley}
	{redhat.com}{aph}
	{Red Hat UK Ltd.}
	{Cambridge, UK}

\member{Dr.\ Ulrich Hoffmann}
	{forth-ev.de}{uho}
	{FH Wedel}
	{Wedel, Germany}

\member{Simon Kaphahn}
	{gmx.de}{Simon\_K99}
	{Independent Member}
	{Munich, Germany}

\member{Dr.\ Peter Knaggs (Editor)}
	{bcs.org.uk}{pjk}
	{Independent Member}
	{Trowbridge, UK}

\member{Bernd Paysan}
	{gmx.de}{bernd.paysan}
	{Net2o}
	{Munich, Germany}

\member{Stephen Pelc}
	{mpeforth.com}{stephen}
	{MicroProcessor Engineering Ltd.}
	{Southampton, UK}

\member[-]{Dr.\ Willi Stricker}
	{t-online.de}{stricker\_w}
	{Independent Member}
	{Springe, Germany}

\member{Leon Wagner}
	{forth.com}{leon}
	{FORTH, Inc.}
	{Los Angeles, USA}

\member{Gerald Wodni}
	{wodni.at}{gerald}
	{Independent Member}
	{Wien, Austria}
\html{</table>}

The following organizations and individuals have also participated in
this project as committee members. The committee recognizes and
respects their contributions:

\html{<table class="center" width="600px">}
\member{Federico de Ceballos}
	{unican.es}{federico.ceballos}
	{Universidad de Cantabria}
	{Santander, Spain}

\member[+]{Andrew Haley}
	{redhat.com}{aph}
	{Red Hat UK Ltd.}
	{Cambridge, UK}

\member{Dr.\ Bill Stoddart}
	{gmail.com}{w.j.stoddart}
	{Teesside University}
	{Middlesbrough, UK}

\member[+]{Dr.\ Willi Stricker}
	{t-online.de}{stricker\_w}
	{Independent Member}
	{Springe, Germany}

\member{Carsten Strotmann}
	{strotmann.de}{carsten}
	{Independent Member}
	{Neuenkirchen, Germany}
\html{</table>}

The committee would like to thank the following contributors:

\begin{minipage}[t]{.3\linewidth}
John		Hayes \\
Marcel		Hendrix \\
Gerry		Jackson \\
Alex		McDonald \\
\end{minipage}
\hfill
\begin{minipage}[t]{.3\linewidth}
Bruce		McFarling \\
Charles G.	Montgomery \\
Krishna		Myneni \\
\end{minipage}
\hfill
\begin{minipage}[t]{.3\linewidth}
Howerd		Oakford \\
Tim			Partridge \\
Elizabeth	Rather \\
David N.	Williams \\
\end{minipage}